---
title: "CARET_Lab"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Package loading
```{r}
library(caret)
library(ggplot2)
```

Load Data
```{r}
# attach the iris dataset to the environment
data(iris)
# rename the dataset
dataset <- iris
```

Task1: Create a Validation/Training Dataset
You need to split the loaded dataset into two, 80% of which we will use to train our models and 20% that we will hold back as a validation dataset.
Hint: use createDataPartition function
```{r}
# Create a vector of row indices for the training set
trainingIndex <- createDataPartition(iris$Species, p = 0.8, list = FALSE)

# Create the training set
trainingSet <- iris[trainingIndex, ]

# Create the validation set
validationSet <- iris[-trainingIndex, ]
```

Task2: Summarize Dataset
Use skimr library to summarize the dataset
```{r}
library(skimr)
# Summarize the dataset
skim(iris)
```
Task3: split input and output
 It is the time to seperate the input attributes and  the output attributes. call the inputs attributes x and the output attribute (or class) y.
```{r}

# Split the data into input and output attributes
x <- iris[, 1:4]
y <- iris[, 5]
```

Task4: Train Control for Validation Test

We will use 10-fold crossvalidation to estimate accuracy.
```{r}
# Run algorithms using 10-fold cross validation
control <- trainControl(method="cv", number=10)
metric <- "Accuracy"
```

Task5: Model Training
Train 5 different algorithms using 'train' function:

- Linear Discriminant Analysis (LDA)
- Classification and Regression Trees (CART).
- k-Nearest Neighbors (kNN).
- Support Vector Machines (SVM) with a linear kernel.
- Random Forest (RF)

```{r}
# Train the LDA model
ldaModel <- train(y ~ ., data = x, method = "lda", trControl = control, metric = metric)

# Train the CART model
cartModel <- train(y ~ ., data = x, method = "rpart", trControl = control, metric = metric)

# Train the kNN model
knnModel <- train(y ~ ., data = x, method = "knn", trControl = control, metric = metric)

# Train the SVM model
svmModel <- train(y ~ ., data = x, method = "svmLinear", trControl = control, metric = metric)

# Train the Random Forest model
rfModel <- train(y ~ ., data = x, method = "rf", trControl = control, metric = metric)

```
